Size matters: Divergent morphology in elliptical bifaces from sites in the American Southeast articulates with distinct local reduction practices

### 2022-11-03 supp/index.Rmd
<edit index>

The pattern that emerges is one of differential retouch intensity, where the isometric reduction strategies of one group (Millsap Cache) may correlate with differing retouch practices that articulate with locally abundant raw material (Kay County flint). This stands in contrast to the comparatively minimal retouch intensity of another group (Jowell Farm) where local raw material was of an insufficient size, yielding a situation where it was necessary to import raw material from an extralocal source (Edwards chert) in central Texas.

Further, the comparison of morphological diversity by size class demonstrated that elliptical biface shapes in the small size class at Jowell Farm occupy a significantly greater range of morphospace than those from Millsap Cache. This pattern suggests that the comparatively minimal retouch intensity associated with Caddo knappers at Jowell Farm may have been more targeted---potentially due to a lack of abundant local raw material of sufficient size---resulting in a broader range of biface shapes in the small size class. Conversely, these results also suggest that Caddo knappers at Millsap Cache employed a more standardized (specialized?) approach to reduction.

>>>>

The pattern that emerges is one of differential retouch _intensity_, where the isometric reduction strategies of one group (Millsap Cache) may correlate with differing retouch practices that articulate with locally abundant raw material (Kay County flint). This stands in contrast to the comparatively minimal retouch intensity of another group (Jowell Farm) where local raw material was of an insufficient size, yielding a situation where it was necessary to import raw material from an extralocal source in central Texas (Edwards chert).

Further, the comparison of morphological diversity by size class demonstrated that elliptical biface shapes in the small size class at Jowell Farm occupy a significantly greater range of morphospace than those from Millsap Cache. This pattern suggests that the comparatively minimal retouch intensity associated with Caddo knappers at Jowell Farm may have been more _targeted_---potentially due to a lack of abundant local raw material of sufficient size---resulting in a broader range of biface shapes in the small size class. Conversely, these results also suggest that Caddo knappers at Millsap Cache employed a more standardized (specialized?) approach to reduction.
